---
title: "p8105_hw3_lg3239"
author: "Landi Guo"
date: "2022-10-16"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

```{r}
head(instacart)
```
### Data Exploration

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

There are 134 aisles, with fresh vegetables and fresh fruits holding the most items.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

This is a plot that shows the number of items ordered in each aisle, which is ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```


This table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, including the number of times each item is ordered.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  #count the number of each product (n)
  count(product_name) %>% 
  #rank within each group according to n
  mutate(rank = min_rank(desc(n))) %>% 
  #three most popular items
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```
This is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

## Problem 2

The column names corresponding to each activity measurement are transformed to one variable `minute`, representing each minute of a 24-hour day starting at midnight. The values are transformed to `activity_measure`, representing the counts for each minute above. The type of `minute` is in numeric. An additional variable `weekday_vs_weekend` is added, specifying whether each measurement is on weekday or on weekend.

```{r}
acc_data = 
  read_csv("hw3_files/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    starts_with("activity"),
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_measure"
  ) %>%
  mutate(minute = as.integer(minute)) %>%
  mutate(weekday_vs_weekend = if_else(day %in% c("Saturday", "Sunday"), "weekend", "weekday"
  )) %>%
  mutate(day = factor(day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")))
view(acc_data)
```

This dataset contains `r nrow(acc_data)` rows and `r ncol(acc_data)` columns, with each row representing the observation of accelerometer data collected within one minute time frame. Variables include number of the week, number of the day, the day of the week, minute since midnight, accelerometer data, and whether on weekday or weekend. These correspond to `r names(acc_data)`. 

This table shows the total activity measurements for each day of the week. This table is intentionally in untidied format for readers. The days with maximum measurement are on the weekends for the first two weeks. The activity measurements are relatively high during the weekdays for the following three weeks.

```{r}
acc_data %>%
  group_by(week, day) %>%
  summarise(total_activity = sum(activity_measure)) %>%
  pivot_wider(
    names_from = "week",
    values_from = "total_activity"
  ) %>%
  knitr::kable(digits = 2)
```


```{r}
acc_data %>%
  ggplot(aes(x = minute, y = activity_measure)) +
  geom_point(aes(color = day), alpha = .5) +
  labs(
    title = "24-hour activity time courses for each day",
    x = "Minute in a day",
    y = "Activity counts"
  ) 
```

